### 2023-04-10 paper-thesis/chapters/01-introduction.tex
Aron's feedback

These contributions are important for two reasons. Firstly from the perspective of a Spoofax researcher, developing a language with a complex type system in Statix tests the boundaries of what Statix can do, and how Statix could be further improved to support these usecases better. On the other hand, from the perspective of a researcher in dependent type systems, if we can show that Statix provides a good way to quickly implement a dependently typed language, then Spoofax could help with rapid prototyping of dependently typed languages.

>>>>

These contributions are important for two reasons. Firstly from the perspective of a Spoofax researcher, developing a language with a complex type system in Statix tests the boundaries of what Statix can do, and how Statix could be further improved to support these use-cases better. On the other hand, from the perspective of a researcher in dependent type systems, if we can show that Statix provides a good way to quickly implement a dependently typed language, then Spoofax could help with rapid prototyping of dependently typed languages.

### 2023-04-10 paper-thesis/chapters/02-background-spoofax.tex
First part of feedback from Jesper
Aron's feedback

The Spoofax Language Workbench~\cite{spoofax} is a platform to develop domain-specific and
general-purpose programming languages. Each aspect of the language is defined in one the meta-languages, each with their own purpose. From these specifications Spoofax generates a parser, type checker and other useful tools. The meta-languages that are used in this thesis are discussed in the sections below:

@@

The constraints are generated by predicates. A predicate defines a relation on terms and scopes, it is of the form $\mathtt{p(t_1, ..., t_n) :-\,\,c_1, ..., c_n}$. Each constraint $c_1...c_n$ is another predicate or an equality constraint. Below is an example that uses these mechanics to check that an expression does not contain any \verb|Muls|. Note that \verb|noMuls| is not defined for the \verb|Mul| constructor, so it will fail if it encounters one.
\begin{lstlisting}
noMuls(Add(e1, e2)) :- noMuls(e1), noMuls(e2).
noMuls(X()).
\end{lstlisting}

@@

Statix also has scope graphs. Scope graphs are used to represent a programs' name and type information~\cite{scopes_as_types}~\cite{nameres}. A scope graph is a graph that consists of the following components:

>>>>

The Spoofax Language Workbench~\cite{spoofax} is a platform to develop domain-specific and
general-purpose programming languages. Each aspect of the language is defined in one the meta-languages, each with its own purpose. From these specifications Spoofax generates a parser, type checker and other useful tools. The meta-languages that are used in this thesis are discussed in the sections below:

@@

The constraints are generated by predicates. A predicate defines a relation on terms and scopes. It is of the form $\mathtt{p(t_1, ..., t_n) :-\,\,c_1, ..., c_n}$. Each constraint $c_1...c_n$ is another predicate or an equality constraint. Below is an example that uses these mechanics to check that an expression does not contain any \verb|Muls|. Note that \verb|noMuls| is not defined for the \verb|Mul| constructor, so it will fail if it encounters one.
\begin{lstlisting}
noMuls(Add(e1, e2)) :- noMuls(e1), noMuls(e2).
noMuls(X()).
\end{lstlisting}

@@

Statix also has scope graphs. Scope graphs are used to represent a programs' name and type information~\cite{scopes_as_types, nameres}. A scope graph is a graph that consists of the following components:

### 2023-04-10 paper-thesis/chapters/03-background-dependent-types.tex
Finish Jesper's feedback
Aron's feedback

The biggest advantage of dependent types is that they increase the expressiveness of a type system a lot, for example:
\begin{lstlisting}
append : (A: Set) -> (n : Nat) -> Vec A n -> Vec A n -> Vec A (n + n)
\end{lstlisting}
A \verb|Vec A n| is a list with elements of type \verb|A| that is exactly length \verb|n| (an integer, which is a value!). This append function appends two \verb|Vecs| of equal length, returning a \verb|Vec| of length \verb|n + n|. This is an example of a type-level expression.

>>>>

The biggest advantage of dependent types is that they increase the expressiveness of a type system, for example:
\begin{lstlisting}
append : (A : Type) -> (n : Nat) -> Vec A n -> Vec A n -> Vec A (n + n)
\end{lstlisting}
A \verb|Vec A n| is a list with elements of type \verb|A| that is exactly length \verb|n| (an integer, which is a value!). This append function appends two \verb|Vecs| of equal length, returning a \verb|Vec| of length \verb|n + n|. This is an example of a type-level expression.

### 2023-04-10 paper-thesis/chapters/11-variable-capture.tex
Aron's feedback

\noindent
The same thing happens, the body of the \verb|FnConstruct| is typechecked with a new scope. Note that \verb|Var("T")| in the type of the second \verb|T| is ambiguous, does it refer to the first or second node?

@@

Another solution that was attempted was having a pre-analysis transformation that gives each variable a unique name. This doesn't work for a variety of reasons. First of all, it doesn't actually solve the problem. Names can be duplicated during beta reduction of terms, so we still don't have the guarantee that each variable has a unique name. Furthermore, this changes the names in the AST, so it breaks editor services in the same was as De Bruijn Indices did.

@@

\begin{lstlisting}
context-free sorts ID
lexical sorts RID
context-free syntax
   ID.Syn = RID
signature constructors
  ScopedName : scope * RID -> ID
\end{lstlisting}

>>>>

\noindent
The same thing happens, the body of the \verb|FnConstruct| is type-checked with a new scope. Note that \verb|Var("T")| in the type of the second \verb|T| is ambiguous, does it refer to the first or second node?

@@

Another solution that was attempted was having a pre-analysis transformation that gives each variable a unique name. This doesn't work for a variety of reasons. This doesn't actually solve the problem. Names can be duplicated during beta reduction of terms, so we still don't have the guarantee that each variable has a unique name. Furthermore, unless we keep track of error messages

@@

\begin{lstlisting}
context-free sorts ID
lexical sorts RID
context-free syntax
   ID.Syn = RID
\end{lstlisting}
\begin{lstlisting}
signature constructors
  ScopedName : scope * RID -> ID
\end{lstlisting}

### 2023-04-10 paper-thesis/chapters/12-editor-services.tex
Aron's feedback

When a language is implemented in Spoofax, we give a declarative definition of the language. Spoofax then generates the parser, type checker and compiler for us. An additional benefit of these declarative definitions is that Spoofax can use them to generate editor services, such as \verb|go to definition|, \verb|renaming| and \verb|semantic code completion|.

>>>>

When a language is implemented in Spoofax, we give a declarative definition of the language. Spoofax then generates the parser, type checker and compiler for us. An additional benefit of these declarative definitions is that Spoofax can use them to generate editor services, such as \emph{go to definition}, \emph{renaming} and \emph{semantic code completion}.

### 2023-04-10 paper-thesis/chapters/20-adding-booleans.tex
Aron's feedback

	\inferrule{
	\bhr
	{ \scope{s}{c} }
	{ [] }
	{ \scope{s'}{c'} }
	\\\\ c' \neq \co{BoolTrue}()
	\\ c' \neq \co{BoolFalse}()
} {
	\bhr
	{ \scope{s}{\co{BoolIf}(c, b1, b2)} }
	{ \overline{p} }
	{ \co{rebuild}((s, \co{BoolIf}(c, b1, b2)), \overline{p}) }
}
\end{mathpar}
\caption{Rules for beta head reducing booleans}
\label{fig:bool-rules-beta}
\end{figure}

@@

This allows each feature to be isolated to its own file. If we decide that we don't like a feature after all, we can remove it simply by unimporting the file. This problem is similar to the expression problem\cite{expression_problem}, where we want to extend datatypes and their associated behaviour, which Statix solves very well.

>>>>

	\inferrule{
	\bhr
	{ \scope{s}{c} }
	{ [] }
	{ \scope{s'}{c'} }
	\\\\ c' \neq \co{BoolTrue}()
	\\ c' \neq \co{BoolFalse}()
} {
	\bhr
	{ \scope{s}{\co{BoolIf}(c, b1, b2)} }
	{ \overline{p} }
	{ \co{rebuild}((s, \co{BoolIf}(c', b1, b2)), \overline{p}) }
}
\end{mathpar}
\caption{Rules for beta head reducing booleans}
\label{fig:bool-rules-beta}
\end{figure}

@@

This allows each feature to be isolated to its own file. If we decide that we don't like a feature after all, we can remove it simply by unimporting the file. This problem is similar to the expression problem\cite{expression_problem, extensability}, where we want to extend datatypes and their associated behaviour, which Statix solves partially. Statix does not fulfill the requirement that files can be individually recompiled.

### 2023-04-10 paper-thesis/chapters/21-adding-inference.tex
Aron's feedback

The syntax we chose is that if we want a value to be inferred, we replace it with an underscore. Other languages such as Agda allow for \emph{implicit arguments}, which are arguments that can be left out, which will then be inferred~\cite{agda}~\cite{coq}. This is just syntactic sugar over the \emph{placeholders} we implement.

@@

There are a lot of different algorithms for inference\cite{typeinference}, some algorithms can solve more inference problems than others. One algorithm for unification is \emph{first-order unification} (FOU), where if at any point during type checking we assert that $\beq{e_1}{e_2}$ and either $e_1$ or $e_2$ is a free metavariable, we set the the value free metavariable to be equal to the value of the non-free metavariable. There are some situations in which this approach fails to infer a term, but in most real-world scenarios it works very well. For example, it can infer both programs in the introduction of this chapter, but it fails to infer the following program:

@@

However, we cannot immediately use Statix' built-in first-order unification (which acts in the meta language, Statix) to implement first-order unification in the object language. Ideally when implementing beta equality when we encounter $\beq{e_1}{e_2}$ we would check if $e_1$ is a free variable, but Statix does not allow for querying whether variables are free.

@@

	\item A complex expression \verb|e1| on the left, an infer expression on the right. In this case, we know what top-level constructor of the metavariable should be, but not necessarily the entire constructor (there might be infers in \verb|e1|). We introduce new \verb|Infer| expressions on the left, and call \verb|expectBetaEq| recursively. A simple example for \verb|BoolIf| is below:
	\begin{lstlisting}
expectBetaEq((s1, e1@BoolIf(c1, t1, b1)), (_, Infer(e2))) :-
	e2 == BoolIf(Infer(c2), Infer(t2), Infer(b2)),
	expectBetaEq((s1, e1), (sEmpty(), e2)).
	\end{lstlisting}
	\verb|FnType| and \verb|FnConstruct| work similarly, but we don't only have sub-expressions but also a name to consider. Sadly, this is the first case where we have to approximate. 
	\begin{lstlisting}
expectBetaEq((s1, e1@FnConstruct(arg_name1, arg_type1, body1)), 
	(_, Infer(e2))) :-
	e2 == FnConstruct(arg_name1, Infer(arg_type2), Infer(body2)),
	expectBetaEq_((s1, e1), (sEmpty(), e2)).
	\end{lstlisting}
	\label{inf-case2}
	Ideally, we would generate a new name iff \verb|e2| is a unresolved metavariable, otherwise using the already generated name. We can't query whether \verb|e2| is free, so as a best-attempt we always assume that \verb|e2| is free and generate a new name, this will fail in some cases. We will discuss this in the next section.

@@

The approximations are not complete: There are situations where AFOU fails to infer a program that FOU can infer. However, these programs are surprisingly uncommon considering how rough the approximations are. They mostly fail when we attempt to infer function values, and two functions are inferred that are beta-equal, and they are beta-equal but not identical. An example of a program where these approximation fails is:

>>>>

The syntax we chose is that if we want a value to be inferred, we replace it with an underscore. Other languages such as Agda allow for \emph{implicit arguments}, which are arguments that can be left out, which will then be inferred~\cite{agda, coq}. This is just syntactic sugar over the \emph{placeholders} we implement.

@@

There are a lot of different algorithms for inference~\cite{typeinference}, some algorithms can solve more inference problems than others. One algorithm for unification is \emph{first-order unification} (FOU), where if at any point during type checking we assert that $\beq{e_1}{e_2}$ and either $e_1$ or $e_2$ is a free metavariable, we set the the value free metavariable to be equal to the value of the non-free metavariable. There are some situations in which this approach fails to infer a term, but in most real-world scenarios it works very well. For example, it can infer both programs in the introduction of this chapter, but it fails to infer the following program:

@@

However, we cannot immediately use Statix' built-in first-order unification (which acts in the meta language, Statix) to implement first-order unification in the object language. Ideally when implementing beta equality when we encounter $\beq{e_1}{e_2}$ we would check if $e_1$ is a free variable, but Statix does not allow for querying whether variables are free. The reason that Statix does not allow this is that it is non-trivial to guarantee confluence if this feature is added to the language.

@@

	\item A complex expression \verb|e1| on the left, an infer expression on the right. In this case, we know what top-level constructor of the metavariable should be, but not necessarily the entire constructor (there might be infers in \verb|e1|). We introduce new \verb|Infer| expressions on the left, and call \verb|expectBetaEq| recursively. A simple example for \verb|BoolIf| is below:
	\begin{lstlisting}
expectBetaEq((s1, e1@BoolIf(c1, t1, b1)), (_, Infer(e2))) :-
	e2 == BoolIf(Infer(c2), Infer(t2), Infer(b2)),
	expectBetaEq((s1, e1), (sEmpty(), e2)).
	\end{lstlisting}
	\verb|FnType| and \verb|FnConstruct| work similarly, but we don't only have sub-expressions but also a name to consider. Sadly, this is the first case where we have to approximate. 
	\begin{lstlisting}
expectBetaEq((s1, e1@FnConstruct(arg_name1, arg_type1, body1)), 
	(_, Infer(e2))) :-
	e2 == FnConstruct(arg_name1, Infer(arg_type2), Infer(body2)),
	expectBetaEq_((s1, e1), (sEmpty(), e2)).
	\end{lstlisting}
	\label{inf-case2}
	Ideally, we would generate a new name if \verb|e2| is a unresolved metavariable, otherwise using the already generated name. We can't query whether \verb|e2| is free, so as a best-attempt we always assume that \verb|e2| is free and generate a new name, this will fail in some cases. We will discuss this in the next section.

@@

The approximations are not complete: There are situations where AFOU fails to infer a program that FOU can infer. However, these programs are surprisingly uncommon considering how rough the approximations are. Inference may fail when we attempt to infer function values, and two functions are inferred that are beta-equal, and they are beta-equal but not identical. An example of a program where these approximation fails is:

### 2023-04-10 paper-thesis/chapters/22-adding-datatypes.tex
Aron's feedback

The general type of an eliminator for a data type \verb|N| is:
\begin{lstlisting}
parameters 
-> P: (indices -> v: N params indices -> Type) 
-> methods
-> indices
-> v: N params indices
-> P indices v
\end{lstlisting}
The meaning of all the arguments is:
\begin{itemize}
	\item \verb|parameters| are the parameters of the data type we want to eliminate, since these are constant among all constructors those should be specified at the start.
	\item \verb|P| is the type that this eliminator will return. It may depend on the value \verb|v| that is eliminated.
	\item \verb|methods| is for each of the constructors of the datatype, a function that eliminates it. To eliminate a constructor of the form \verb|C : args -> N params indices| it requires a function \verb|args -> P indices (C pars args)|. For recursive datatypes, a previous case is generated, such as the \verb|P x| in the eliminator of \verb|Nat| above.
	\item \verb|indices| are the indices of the value that is to be eliminated.
	\item \verb|v| is the value that is to be eliminated.
	\item Finally, \verb|P indices v| is the result of the elimination.
\end{itemize}

@@

	\item Create a chain of scopes starting in \verb|s2| that will contain a scope for each index. We will type-check each next index using the previous index' scope, so that one index can depend on parameters and previous indices. Call the end of this chain \verb|s3|. The scope \verb|s3| is not used in the rest of this process, but creating this chain is required to check the types of the indices.

>>>>

The general type of an eliminator for a data type \verb|N| is:
\begin{lstlisting}
params 
-> P: (indices -> v: N params indices -> Type) 
-> methods
-> indices
-> v: N params indices
-> P indices v
\end{lstlisting}
The meaning of all the arguments is:
\begin{itemize}
	\item \verb|params| are the parameters of the data type we want to eliminate, since these are constant among all constructors those should be specified at the start.
	\item \verb|P| is the type that this eliminator will return. It may depend on the value \verb|v| that is eliminated.
	\item \verb|methods| is for each of the constructors of the datatype, a function that eliminates it. To eliminate a constructor of the form \verb|C : args -> N params indices| it requires a function \verb|args -> P indices (C params args)|. For recursive datatypes, a previous case is generated, such as the \verb|P x| in the eliminator of \verb|Nat| above.
	\item \verb|indices| are the indices of the value that is to be eliminated.
	\item \verb|v| is the value that is to be eliminated.
	\item Finally, \verb|P indices v| is the result of the elimination.
\end{itemize}

@@

	\item Create a chain of scopes starting in \verb|s2| that will contain a scope for each index. We will type-check each next index using the previous index' scope, so that one index can depend on parameters and previous indices. Call the end of this chain \verb|s3|. The scope \verb|s3| is not used in the rest of this process (since constructors cannot refer to indices), but creating this chain is required to check the types of the indices.

### 2023-04-10 paper-thesis/chapters/24-backend.tex
Third part of feedback from Jesper
Aron's feedback

Luckily, Stratego has a strategy built in that runs a Statix predicate on a term: \verb|stx-evaluate|. Using this strategy, the actual implementation of our interpreter in Stratego is just one line: Match an expression \verb|e| and run the functional predicate \verb|interpret| from the \verb|main.stx| file on it.

@@

Another disadvantage is that Stratego is specifically optimized for these transformations, for example as discussed in "Optimising First-Class Pattern Match Compilation"~\cite{fcpmc}. Therefore, the resulting interpreter will be slower when implemented in Statix. This is however just a constant factor overhead, it does not affect the time complexity of the algorithm.

@@

Types require a bit more attention. Since our language has no way to match on types, types can actually be completely eliminated during type checking. The easiest way to do this is to compile all types to an arbitrary value (we chose the string "TYPE"), since the value is irrelevant anyways. A more efficient solution could be to implement type erasure~\cite[Section 23.6]{tapl}, where type-level functions are completely removed. We decided against this as the goal of this section is to show that we can compile the language with ease, not to write an efficient compiler. This optimization was also not implemented in the Agda to Scheme compiler.

>>>>

Luckily, Statix exposes an API to Stratego that can run a Statix predicate on a term: \verb|stx-evaluate|. Using this strategy, the actual implementation of our interpreter in Stratego is just one line: Match an expression \verb|e| and run the functional predicate \verb|interpret| from the \verb|main.stx| file on it.

@@

Another disadvantage is that Stratego is specifically optimized for these transformations, for example as discussed in "Optimising First-Class Pattern Match Compilation"~\cite{fcpmc}. Therefore, the resulting interpreter will be slower when implemented in Statix.

@@

Types require a bit more attention. Since our language has no way to match on types, types can actually be completely eliminated during compilation. The easiest way to do this is to compile all types to an arbitrary value (we chose the string "TYPE"), since the value is irrelevant anyways. A more efficient solution could be to implement type erasure~\cite[Section 23.6]{tapl}, where type-level functions are completely removed. We decided against this as the goal of this section is to show that we can compile the language with ease, not to write an efficient compiler. This optimization was also not implemented in the Agda to Scheme compiler.

### 2023-04-10 paper-thesis/chapters/40-discussing-statix.tex
Third part of feedback from Jesper
Aron's feedback

\begin{enumerate}
	\item Allow specifying reduction rules, and then implement equality under reduction in Statix. This feature was discussed extensively in chapter \ref{ch:comp-lambdapi}.

@@

	isStrictlyPositive : (scope * Expr)
	isStrictlyPositive(s, FnType(...)) = ...
	\end{lstlisting}
	The definition of \texttt{isStrictlyPositive} can then use all the metavariables of \\ \texttt{DataTypeDecl} without having to explicitly pass them.

	\item Generic constraints. For example, we created two separate relations for reversing lists of different types, making these one generic function would've been better. This was the only case where this feature would've been useful for this thesis, but it may be more useful for other work.
\end{enumerate}

@@

	\item The SPT testing DSL works very well, it makes writing integration tests very easy. However, support for unit tests is still lacking. The to create unit tests is using Statix tests, allowing assertions on specific relations. But these tests are not integrated with SPT tests (they don't even reside in the test folder) and there is no way to run all Statix tests, to ensure no regressions happened. Ideally there would be one button that runs all SPT and Statix tests.

	\item Debugging Statix is still quite tedious. When we have a constraint that does not do what we expect it to do (ie. we have a failing Statix test), one either has to stare at its definition until they figure out the problem, or start unfolding the definitions of the constraints into the Statix test, finding out which part of the definition is misbehaving and repeating the unfolding. Having support for a debugger would help tremendously with this.
\end{enumerate}

>>>>

\begin{enumerate}
	\item Allow specifying reduction rules, and then implement equality under reduction in Statix. This feature was discussed extensively in chapter \ref{ch:comp-lambdapi}. This would require extensive changes to the core language of Statix, so this is challenging to implement.

@@

	isStrictlyPositive : (scope * Expr)
	isStrictlyPositive(s, FnType(...)) = ...
	\end{lstlisting}
	The definition of \texttt{isStrictlyPositive} can then use all the metavariables of \\ \texttt{DataTypeDecl} without having to explicitly pass them. This feature could be desugared by moving the relation to the global scope and explicitly passing the metavariables of the parent constraint when it is used.

	\item Generic constraints. For example, we created two separate relations for reversing lists of different types, making these one generic function would have been better. This was the only case where this feature would have been useful for this thesis, but it may be more useful for other work. This could be implemented by monomorphization or by changing the core language. 
\end{enumerate}

@@

	\item The SPT testing DSL works very well, it makes writing integration tests very easy. However, support for unit tests is still lacking. The way to create unit tests for the type checker is using Statix tests, allowing assertions on specific relations. But these tests are not integrated with SPT tests (they don't even reside in the test folder) and there is no way to run all Statix tests, to ensure no regressions happened. Ideally there would be one button that runs all SPT and Statix tests.

	\item Debugging Statix is still quite tedious. When we have a constraint that does not do what we expect it to do (ie. we have a failing Statix test), one either has to stare at its definition until they figure out the problem, or start unfolding the definitions of the constraints into the Statix test, finding out which part of the definition is misbehaving and repeating the unfolding. Having support for a debugger would help tremendously with this.

	The way that this debugger could work is that you could start by entering a certain constraint \verb|typeCheck(...)|, the debugger would then show the current value of each metavariable in scope, and each constraint that is yet to be expanded. Since there is no fixed evaluation order in Statix, we could even allow the programmer to choose which constraint is evaluated next, which then generates more constraints.
\end{enumerate}

### 2023-04-10 paper-thesis/chapters/98-relatedwork.tex
Aron's feedback

The implementation in this paper requires performing substitutions in types immediately, as types don't have a corresponding scope to store substitutions. Van Antwerpen et al.~\cite[sect 2.5]{scopes_as_types}~\cite{nameres} present an implementation of System F in Statix that does lazy substitutions, by using scopes as types. Each type is represented by a scope, and these scopes have edges to other scopes in case of quantifications and substitutions. Our Statix implementation does lazy substitutions in the expressions of our language, but not in the types. This is a minor compromise since types tend to stay quite small, but it is still a compromise. Further research is needed to see if this approach could also apply to dependently typed languages, where types can contain arbitrary terms.

>>>>

The implementation in this paper requires performing substitutions in types immediately, as types don't have a corresponding scope to store substitutions. Van Antwerpen et al.~\cite[sect 2.5]{scopes_as_types, nameres} present an implementation of System F in Statix that does lazy substitutions, by using scopes as types. Each type is represented by a scope, and these scopes have edges to other scopes in case of quantifications and substitutions. Our Statix implementation does lazy substitutions in the expressions of our language, but not in the types. This is a minor compromise since types tend to stay quite small, but it is still a compromise. Further research is needed to see if this approach could also apply to dependently typed languages, where types can contain arbitrary terms.

### 2023-04-10 paper-thesis/chapters/99-conclusion.tex
Aron's feedback

	\item Another feature of Agda is implementing dependent pattern matching, as described by Coquand~\cite{patternmatching}. Dependent pattern matching is complex because it requires interleaved type checking and desugaring, which is problematic since desugaring is usually done in Stratego, while type checking is done in Statix. The solution may be to do the desugaring in Statix instead, but at some point this is not ideal, since Statix is not meant for this. More research is needed on a good solution for this problem.

@@

	\item Agda has holes, which allows the user of the language to build proofs interactively. One of these interactions is proof search, when triggered Agda will try to find a term that fits the expected type of a hole. The interactive part of this was discussed in chapter~\ref{ch:relatedwork}, this is not yet supported by Spoofax. After support for this is added, a proof search algorithm could be implemented in Stratego, such as the structurally recursive unification algorithm described by McBride~\cite{proofsearch}. Running the algorithm may require running type checking code, so here we encounter the interleaving problem again. More research is needed on whether proof search can be implemented in Spoofax.

>>>>

	\item Another feature of Agda is dependent pattern matching, as described by Coquand~\cite{patternmatching}. Dependent pattern matching is complex because it requires interleaved type checking and desugaring, which is problematic since desugaring is usually done in Stratego, while type checking is done in Statix. The solution may be to do the desugaring in Statix instead, but at some point this is not ideal, since Statix is not meant for this. More research is needed on a good solution for this problem.

@@

	\item Agda has holes, which allows the user of the language to build proofs interactively. One of these interactions is proof search, when triggered Agda will try to find a term that fits the expected type of a hole. The interactive part of this was discussed in chapter~\ref{ch:relatedwork}, this is not yet supported by Spoofax. After support for this is added, a proof search algorithm could be implemented in Stratego, such as the structurally recursive unification algorithm described by McBride~\cite{proofsearch}. The holes system that was introduced for semantic code completion~(chapter~\ref{chap:editor-services}) could be re-used for this purpose. Running the algorithm may require running type checking code, so here we encounter the interleaving problem again. More research is needed on whether proof search can be implemented in Spoofax.
